### 2019-01-07 ms.tex
Updated version on Overleaf

%%
%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%

% These are my private definitions

\newcommand\LL{\mbox{$\:\lambda\lambda $ }}
\newcommand\lam{\mbox{$\:\lambda\ $}}
\newcommand\lamlam{\mbox{$\:\lambda\lambda $ }}
\newcommand\ha{{H$\alpha$}}
\newcommand\hb{{H$\beta$}}

%Needs trailing \ except at end of sentence
\newcommand\BT{B_T}
\newcommand\MBT{M_{B_T}}
\newcommand\Mstar{M^*}
\newcommand\Lstar{L^*}
\newcommand\Mpc{\:\rm{Mpc}}
\newcommand\kpc{\:\rm{kpc}}
\newcommand\pc{\:\rm{pc}}
\newcommand\kms{\:\rm{km\,s^{-1}}}
\newcommand\rfid{r_{\rm{fid}}}

\newcommand\LUM{\:{\rm erg\:s^{-1}}}
\newcommand\FLUX{\:{\rm erg\:cm^{-2}\:s^{-1}}}
\newcommand\FLUXARCSEC{\:{\rm erg\:cm^{-2}\:s^{-1}\:arcsec^{-2}}}
\newcommand\COUNTS{\:{\rm cts\:s^{-1}}}
\newcommand\CPM{\:{\rm s^{-1}\:arcmin^{-2}}}
\newcommand\VEL{\:{\rm km\:s^{-1}}}
\newcommand\ETAL{{\it et\:al.}}
\newcommand\etal{{\it et\:al.}}

\newcommand\OIGS{\:{\rm ergs\:cm^{-2}\:s^{-1}\:\AA^{-1}}}
%\newcommand\LA{Lyman\thinspace$\alpha$}
%\newcommand\LB{Lyman\thinspace$\beta$}
%\newcommand\LG{Lyman\thinspace$\gamma$}
%def\LD{Lyman\thinspace$\delta$}

%Use Roman numbers to designate ionization state and first two digits
%of wavelength in Ang. where necessary

\newcommand\OiL{[\ion{O}{1}] $\lambda 6300$}
\newcommand\OiiL{[\ion{O}{2}] $\lambda\lambda 3727, 3729$}
\newcommand\oiiL{[\ion{O}{2}] $\lambda 3727$}
\newcommand\OiiiL{[\ion{O}{3}] $\lambda\lambda 4959,5007$}
\newcommand\SiiL{[\ion{S}{2}] $\lambda\lambda 6717, 6731$}
\newcommand\NiiL{[\ion{N}{2}] $\lambda\lambda 6548, 6583$}
\newcommand\OiLL{[\ion{O}{1}] $\lambda\lambda 6300, 6363$}

\newcommand\sii{[\ion{S}{2}]}
\newcommand\nii{[\ion{N}{2}]}
\newcommand\oi{[\ion{O}{1}]}
\newcommand\oii{[\ion{O}{2}]}
\newcommand\oiii{[\ion{O}{3}]}
\newcommand\feii{[\ion{Fe}{2}]}
\newcommand\hii{\ion{H}{2}}
\newcommand\hi{\ion{H}{1}}
\newcommand\perpix{\: \rm pixel^{-1}}
\newcommand\gal{NGC\,6946}

%private definitions end here

%\begin{document}

\documentclass[modern]{aastex62}

%%
%% using aastex version 6.2

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex61}
%% 
%% where the arguement options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}

%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
% \received{July 1, 2016}
%\revised{September 27, 2016}
%\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
%\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{SNRs in NGC~6946}
\shortauthors{Long, Winkler, \& Blair}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.
%\include{pfw_defns}

\usepackage{natbib}
\usepackage{comment}
\usepackage{todonotes}
\begin{document}
%\slugcomment{Draft \today}

\title{An New Search for Supernova Remnants in NGC\,6946}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.1 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.1 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Knox S. Long}
\email{long@stsci.edu}

\author[0000-0002-4134-864X]{Knox S. Long}
\affil{Space Telescope Science Institute,
3700 San Martin Drive,
Baltimore MD 21218, USA; long@stsci.edu}
\affil{Eureka Scientific, Inc.
2452 Delmer Street, Suite 100,
Oakland, CA 94602-3017}

\author[0000-0001-6311-277X]{P. Frank Winkler}
\affiliation{Department of Physics, Middlebury College, Middlebury, VT, 05753; 
winkler@middlebury.edu}

\author[0000-0003-2379-6518]{William P. Blair}
\affiliation{The Henry A. Rowland Department of Physics and Astronomy, 
Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD, 21218; 
wblair@jhu.edu}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

The relatively nearby spiral galaxy NGC~6946 is one of the most actively star forming galaxies in the local Universe.  Ten supernovae have been observed since 1917, and hence NGC~6946 surely contains a large number of supernova remnants (SNRs).  
Here we report a new optical search for these SNRs in NGC~6946 using narrow-band images obtained with the WIYN telescope to 
identify 147 emission nebulae 
as likely SNRs,  based on elevated \sii:\ha\ ratios compared to \hii\ regions.  We have obtained spectra of 102 of these emission nebulae with Gemini North--GMOS; of these, 89 have \sii:\ha\ ratios greater than 0.4.  All but one of the 27 SNR candidates identified earlier by Matonick \& Fesen (1997) are contained in our sample.  However, there is very little overlap in our sample between the SNR candidates identified by Bruursema et al. (2014) from \feii\ imagery or by Lacey et al. (2001) from radio data.  Very few of the SNR candidates are X-ray sources, unlike the situation in some other galaxies such as M33 and M83, possibly due to  more reddening along the line of sight to NGC~6946.  The emission line ratios, e.g., \nii:\ha\ of the candidates in NGC~6946 are fairly typical of those observed in SNR samples from other galaxies with comparable metallicity.  
Two of the historical supernovae were detected as young optical SNRs showing high velocity ejecta. None of the other candidates observed spectroscopically show evidence of anomalous abundances or evidence of significant velocity broadening in our low resolution spectra.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{galaxies: individual (NGC~6946) -- galaxies: ISM  -- supernova remnants}

\todo[inline]{WPB: This will likely have to get shortened further before submission.}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

\gal\ is a nearby \cite[$6.72\pm 0.15$ Mpc,][]{tikhonov14}, nearly face-on \cite[$i = 32.6\degr$,][]{deblok08} galaxy with four flocculent spiral arms.  The galaxy is currently undergoing a major starburst, and as a result it has been the site of ten historical supernovae (SNe) since 1917, the most of any known galaxy. According to \citet{jarrett13} the total star formation rate is $3.2\, M_\sun\,{\rm yr}^{-1}$, a high rate that is thought to be bar-driven.  A total of 121 bubbles, probably created by stellar winds and multiple SNe at the star-formation sites, have been identified in \hi\ gas that extends well outside the bright portions of the optical galaxy \citep{boomsma08}.  %The high SFR in \gal\ is thought to be bar-driven.   
Given these properties, one expects that a large number of supernova remnants (SNRs) should be present in \gal, since SNRs tend to remain visible for thousands of years.

Optically, SNRs are usually identified on the basis of high \sii:\ha\ ratios compared to \hii\ regions.  In bright \hii\ regions, most S is found in the form of S$^{++}$ or S$^{+++}$, and as a result the \sii:\ha\ ratios are typically 0.1 or smaller.  In SNRs, where emission is driven by impulsive heating from shock waves, S$^+$ is found in an extended recombination and cooling zone behind the shock, and the \sii:\ha\ ratios are typically $\gtrsim0.4$.  The diagnostic can become less deterministic at lower surface brightness, as recently discussed by \cite{long18} for the case of M33.

The first optical search for SNRs in \gal\ was made by \cite{matonick97}, hereafter MF97, who used interference filter imagery to identify 27 emission nebulae with \sii:\ha\ ratios $\geq$ 0.45 as SNRs.  One of these sources, MF97-16, was, later associated with the ultraluminous black hole X-ray binary NGC~6946 X-1 \citep{roberts03}. Though very rare, such ultra-luminous X-ray sources (ULXs) have  the hard X-ray spectra that produce line ratios in the surrounding circumstellar ISM that resemble those expected from SNRs. To our knowledge, no other optical searches for SNRs in NGC~6946 exist, nor have spectra of the remaining MF97 objects ever been reported.

\hii\ regions and SNRs also radiate at radio wavelengths.  SNRs typically have steep non-thermal radio spectral indices arising from synchrotron radiation, whereas \hii\ regions have relatively flat thermal spectra arising from free-free and recombination radiation. \todo{KSL to check previous assertion}  At radio wavelengths, SNRs in nearby galaxies are generally identified as non-thermal radio sources associated with optical \ha\ emission.  (This association criterion arises because many background radio sources---typically active galaxy nuclei (AGNs)---also have non-thermal spectra, and like AGNs, the SNRs are generally unresolved in radio images of galaxies beyond the Local Group.)  In the case of \gal,  \cite{lacey01} identified 35 radio-selected SNR candidates on the basis of these criteria, none of which overlap with the MF97 list of optical candidates.  Their radio candidates were located in the brightest parts of the spiral arms of \gal; Lacey \& Duric suggested that the reason the two lists do not overlap is primarily due to selection effects.

Here we discuss a new, more sensitive optical search for SNRs in \gal\ in which we identify a total of 147 SNR candidates using interference filter imagery.  We also discuss spectroscopic observations of 102 of these candidates, which we use to verify the ratios obtained from the imaging and to characterize other characteristics of our new optical sample.  The paper is organized as follows:  Section 2 describes both our imaging and spectroscopy observations,  presents our much-expanded catalog of optical SNR candidates, and the results of our spectroscopy that has confirmed many of these.  Section 3 discusses these results in the context of both \gal\ and other spiral galaxies, and Section 4 provides a brief summary of our results.

\section{Observations and Data Reduction \label{sec:observations}}
\subsection{Imaging and Catalog of SNR Candidates}

We carried out narrow-band imaging observations of \gal\ from the 3.5m WIYN telescope and MiniMosaic imager on Kitt Peak on the nights of 2011 June 26-28 (UT).  The so-called ``Minimo'' was mounted at the f/6.3 Nasmyth port and consisted of a pair of $2048\times4096$ SITe chips, with a field  9\farcm 6 square at a scale of 0\farcs 14 pixel$^{-1}$.  We used interference filters that pass lines of \ha, \sii \lamlam6716,6731, and \oiii \lam 5007, plus red and green narrow-band continuum filters so we could subtract the stars and produce pure emission-line images. Frames in each filter were dithered to enhance removal of cosmic rays and bad pixels.  Further observational details are given in Table 1.  It is noteworthy that the \ha\ filter has quite a narrow bandwidth, 27\ \AA\ FWHM. Its transmission is 69\% at the rest wavelength of \ha, but only 11\% at 6548\ \AA\ and 16\% at 6583\ \AA; hence, the \nii\ lines are greatly attenuated relative to \ha.\footnote{The recessional velocity of \gal\ is only $40 \kms$, so the lines are redshifted by $\lesssim 1$ \AA.}  This facilitates identification of SNRs based on their image-derived \sii:\ha\ ratios.  Seeing throughout this run was about $1\arcsec \approx 32$ pc at the distance of \gal.  While sufficient for identifying SNRs in all but the most crowded regions, this resolution could provide only crude diameter estimates and only for the larger objects.

We used standard IRAF\footnote{IRAF is distributed by the National Optical Astronomy Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.} techniques for processing the images, including overscan correction, bias subtraction, and flat-fielding using dome flats.  Procedures in the IRAF {\texttt mscred} package were then used to combine the data from the individual chips into a mosaic image for each frame, assigned a WCS for each using stars from the USNOB1 catalog (WPB: ref needed?).  We then stacked  all the images by filter onto an arbitrary standard coordinate system with a scale of 0\farcs 20 pixel$^{-1}$ and scaled and subtracted the continuum images from the emission-line ones (red from \ha\ and \sii; green from \oiii) to give pure emission-line images with most of the stars and galaxy background light removed.  Finally, we used observations of several spectrophotometric standard stars from the catalog of \citet{massey88} to flux calibrate the emission-line images.  Figure \ref{fig_overview} shows a color version of the final images (R = \ha, G = \sii, B = \oiii).

\todo[inline]{WPB: For the Overview Figure, I like the version that shows the galaxy light (currently Fig. 2).  Then in addition to the overview figure, we should show an enlargement of a smaller region in a multi-panel format, showing the subtracted data.  Perhaps this could be done in conjunction with the following paragraph:  Show a small region with several \hii\ regions and SNRs, where the color panel shows the SNRs standing out in different color.  Could include ratio panels as well.}

To select SNR candidates, we used the DS9 display program to show the continuum-subtracted WIYN images in all three emission lines, as well as \sii:\ha\  and \oiii:\ha\ ratio images, and a continuum image (to distinguish stars  or stellar subtraction residuals from point-like nebulae).   We then visually inspected these to select SNR candidates based on a high  \sii:\ha\ ratio.  The initial inspection was carried out by Middlebury undergraduate Marc DeLaney; subsequently two of us (WPB and PFW) conducted independent inspections, and then we conferred to agree on a consensus list of 147 candidates, including the 27 from MF97.   In addition, we selected 51 objects with relatively high \oiii:\ha\ ratios, in an attempt to identify young, ejecta-dominated SNRs such as Cas A in the Galaxy. With ten SNe in the last century, one might expect a number of young, ejecta-dominated SNRs to be present.  Though none of the O-selected candidates had ratios as extreme as seen for Cas A or 1E0102-72.3 in the Small Magellanic Cloud, we nevertheless selected some of these for follow-up spectroscopy.  Finally, we inspected the positions of the nine\footnote{As of the 2011 observations there were nine SNe; a tenth SN was recorded in 2017.} historical SNe in \gal\ for evidence of nebular emission, and detected line emission only at the positions of SN1980K and SN2004et, both of which we targeted for follow-up spectra.

\todo[inline]{Add symbols to Fig. 1 overview and reference the Figure again here?  I note below we intend to also indicate those observed spectroscopically with a separate symbol or different color.  Could be tricky. WPB}

The SNR candidates are listed in Right Ascension order in Table \ref{table_candidates}.  For each candidate, we provide (1) a source name, (2) the name of the source in MF97 if any,  (3,4) the position  (J2000), (5) the deprojected galactocentric distance (GCD), (6) the \ha\ flux as derived from the emission line images, (7) the \sii:\ha\ ratio measured from the images, (8) the spectrum we used to confirm the imaging ratios (see below), and (9) whether or not the object has a spectroscopic confirmation that the  \sii:\ha\ ratio is $\ge$ 0.4 (see below).

\subsection{Spectroscopy; Emission-Line Fluxes}

%We used a combination of procedures in the GMOS and apextract packages of IRAF to reduce the spectra.

We used the Gemini Multi-Object Spectrograph (GMOS) on the 8.2m Gemini-North telescope to obtain all the spectra reported here, during queue-scheduled programs in semesters 2014B (program GN-2014B-Q-83) and 2015B (program GN-2015B-Q-91).  
For the 2014B program, we designed six custom masks, each with 20-30 slitlets targeting SNR candidates whose positions we determined from our 2011 WIYN images, together with  short $R$-band pre-images of several \gal\ fields taken with GMOS earlier in 2014 as part of the spectroscopy preparation program.    
We used two additional masks (which we refer to as masks 7 and 8 for simplicity)  for the 2015B program.  Slitlets in one or more of our eight masks were placed on 102 distinct SNR candidates, including ones with a range of sizes, GCDs, and ISM environments (locations in arms and in 
inter-arm regions).  In addition to the SNR candidates, we also placed a number of slitlets on prominent \hii\ regions for comparison purposes in both 2014 and 2015.
Figure \ref{fig_overview} shows all the objects from the sample for which we obtained spectra (red boxes).

For all the spectra we used GMOS-N with the 600 line\,mm$^{-1}$ grating designated G5307 and a GG455 cut-off filter to block second-order spectra.   The detector in both years was a mosaic of three e2v deep-depletion CCD chips, binned by 2 in the spatial direction (for a scale of 0\farcs 146 pixel$^{-1}$) and by 4 in the dispersion direction.  
The dispersion was 1.84 \AA\,  pixel$^{-1}$ (binned), resulting in coverage of the spectral range from at least \hb\ through \sii\lamlam 6716,\,6731 for virtually all the objects.\footnote{The detailed wavelength coverage for individual objects naturally varied with slitlet position in the dispersion direction.}  Our  masks had  slitlet  widths from 1\farcs 25 to 1\farcs 75, with wider slits used for the larger objects, and lengths of 6\arcsec\ or longer to permit local background and sky subtraction.

With each mask, we took spectra at three or four slightly different grating tilts, to cover wavelength gaps between chips and to gain somewhat more total spectral range.\footnote{An exception was mask 6, done late in the 2014B semester, for which our full set of planned observations were never completed.}  At each wavelength setting, we obtained two or more identical exposures to minimize the effects of cosmic rays.
For calibration, we programmed quartz flats and CuAr arc frames immediately before or after the science exposures with each mask and grating setting.  A journal of all the science observations from both 2014 and 2015 appears in Table \ref{obs_log}.

%\todo[inline]{WPB: Next paragraph: Frank, do we need to say something about how the separate grating offset exposures were shifted and co-added prior to obtaining the final 2-D spectra for each slitlet?}

The data were processed using standard procedures  from the {\tt gemini} package in IRAF  for bias subtraction, flat-fielding, wavelength calibration, and combination of spectra with different grating tilts to provide the final results.  
Flux calibration was based on baseline GMOS observations of a few spectrophotometric standard stars, carried out in the same semester as part of standard GMOS operations.  
%  IRAF footnote already given above, so I commented this out.  WPB
%\footnote{IRAF is distributed by the National Optical Astronomy Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.}  
During the processing, the 2-D spectra from different slitlets were separated to give individual 2-D spectra from each slitlet.   We examined each of these individually and selected the object region, as well as one or more sky background regions,  stripped out 1-D spectra of each, and then subtracted the sky spectrum from the corresponding object to obtain the final background-subtracted object spectra.
Many of the objects are located in regions with bright surrounding galactic background (both continuum and emission lines) from NGC\,6946, so the selection of a representative local background was done on a best-effort basis.  
We then performed  fits to obtain emission line fluxes  from the spectra, assuming Gaussian profiles.  We performed fits for various lines and line complexes:  \hb\ alone,  the \oiii\ doublet,  the \oi\ doublet,  the \ha-\nii\ region, and   the \sii\ doublet. For the fits, we assumed that the background varied linearly with wavelength around each line, and that the FWHM of all lines in each complex was the same.

\todo[inline]{It would probably be good to include a figure showing a few example spectra, as in M83 paper, Fig 3.}

Table \ref{table_snr_spec}  lists the information we obtained for the SNR candidates for which we obtained spectra.  Specifically we list (1) the source name, (2) the extracted  \ha\ flux, (3-9) ratios of various emission lines to \ha\ [taken to be 300], (10) the total \sii:\ha\ ratio and (11) the measured FWHM of the lines in the \ha-\nii\ complex.  For doublets, where the line ratio is constrained by atomic physics, that is \oiii, \oi, and \nii, we have listed only the stronger line.  We visually inspected all of the spectra and the fits to them; values which we judge to be more uncertain are indicated with tildes.  No allowance has been made for additional errors associated with difficulties in background subtraction.  A number of the objects were observed several times, sometimes with different slit orientations.  In these cases, we used the spectrum which we judged to be the most accurate and report it for reference in the `Spectrum' column in Table \ref{table_candidates}, where (for example) 05.18 should be interpreted as `mask 5 slitlet number 18.'

\section{Analysis and Discussion}

We identified 147 candidate SNRs from the WIYN interference images; of these we obtained spectra of 102. The spectra were needed to improve our confidence that the \sii:\ha\ ratios of the emission nebulae were indeed high and to begin to characterize the SNRs using accurate line ratios.  The spectroscopic ratios are shown as a function of GCD in the left panel of  Fig.\ \ref{fig_s2_ha}. \hii\ regions and \oiii-selected objects are shown in blue while nominal \sii:\ha\ candidates from imaging are shown in red.   Conventionally, emission nebulae are identified as SNRs optically, if the \sii:\ha\ ratio exceeds 0.4; 89 of the 102 SNRs with spectra satisfy this criterion, and so we regard these as confirmed SNRs (and they have been listed as such in Table \ref{table_candidates}. Clearly, given uncertainties in derived line ratios for faint emission line objects, a dividing line of 0.4 is somewhat arbitrary, and objects just above (or just below) this ratio should be judged with more context.  For example, slightly under-subtracting contaminating \ha\ emission could readily explain why some candidates ended up below the threshold in the spectroscopic analysis. Likewise, a slight over-subtraction of \ha\ could enhance the \sii:\ha\ ratio derived, potentially pushing some objects above the threshold.  The observed tendency to see higher \sii:\ha\ ratios closer to the nucleus is likely to be at least partly an abundance effect; similar trends are seen in M33 \citep{long18} and in M83 \citep{winkler17}.

There are 45 objects without spectra, so which of these are actually SNRs is uncertain.  That said, as shown in the right panel of Fig.\ \ref{fig_s2_ha}, if anything the spectroscopically-determined ratios tend to be higher than the ratios determined from narrow band imaging, and so most of the objects without spectra are also likely to be SNRs.  The somewhat higher spectroscopic ratios are to be expected, since even with our relatively narrow \ha\ filter , some emission from \nii\ was  also passed.

Of the 27 SNR candidates identified by MF97, all have imaging \sii:\ha\ ratios that exceed 0.4. MF97 obtained spectra of six of their candidates.  We have obtained spectra of 23 of the MF97 objects, including new spectra of four objects for which MF97 had spectra -- MF003, MF021, MF022, and MF026.  All of these indeed have measured spectroscopic ratios that exceed 0.4.

\todo[inline]{We need a more complete discussion of MF97 and our results. MF-016 probably needs several sentences as it is a unique object. Possibly this is something wpb could do.  WPB: possibly, if it can wait until I return.}

As shown in Fig.\ \ref{fig_s2_ratio}, the ratios of [S~II]$\lambda$6717:S~II]$\lambda$6731 for the SNR candidates cluster around the low density limit of 1.4, and from the fact that just about as many objects have non-physical ratios above 1.4 as below, suggests that the all of the objects are close to this limit.  In the absence of SNR diameters, we cannot search for trends with diameter, as were seen in M33 (L18) and M83 \citep{winkler17}. Upcoming {\it Hubble Space Telescope} observations will provide accurate diameter information for many of these objects.

Fig.\ \ref{fig_reddening} shows the observed \hb:\ha\ ratios of the SNR candidates with spectra as a function of GCD. Nearly all of the SNRs show significant reddening, as one would expect since, at 12\degr\ from the Galactic plane, foreground reddening from within our Galaxy along the line of sight is expected to be E(B-V) = 0.29 \citep{schlafly11}.  There is clearly substantial internal and differential reddening  within \gal\ as well, as there is a very evident trend for objects near the center of \gal\ to be more reddened than those more distant from the nucleus.

\todo[inline]{Bill and Frank point out that various other diagrams with GC radius are missing, and in particular \sii:\ha\ ratio vs. \ha\ flux; \sii:\ha\ ratio vs. GCD (fig for this already here)  Check M83 to see if there is a correlation in the spectroscopic ratio with GCD, and if so comment on this.  This is done for S2; I guess we could to it for N2. Gradients: \nii:\ha\ and \sii:\ha\ vs GCD (M83 paper, fig 8) [WPB: Yes, I'd definitely like to see what this looks like.].  ksl: This is now provided, but whether these should be included is open to question.  If they are included some massaging of the plots is needed.  If included, Bill or Frank should provide words, and comparison to other galaxies}

Some of the line ratios show indications of a metallicity gradient as evidenced in Fig. \ref{fig_metal_ratios}.  However, unlike the case in M33, what is seen in both the \nii:\ha\ and \sii:\ha\ ratio is that the upper envelope of both plots decreases with radius, but there are examples of SNRs with low ratios at all galactocentric distances.

\todo[inline]{Frank and Bill have suggested that we do the comparison to models figures in the same way we did for M33 in L18, which includes plots not only of the models but also of where these ratios lie in other galaxies. See Figures 10 and 11 of L18. Bill also thinks the ratios are less than in M83 and wonders why that might be }

Line ratios in spectra of SNRs are expected to vary both as a function of shock conditions and metallicity.  To see where the SNRs in \gal\ lie, we have compared the line ratios calculated from models by \cite{allen08} using the MAPPINGS III code for a range of shock velocities (100 - 1000 $\VEL$) and pre-shock magnetic fields (10$^{-4}$ - 10 $\mu$G).  The results are shown in Fig. \ref{fig_model}.  The results both for the ratio of [O~III]5007:\hb\ and for \sii:\ha\ fall squarely into the region of the solar metallicity models, as one would expect for \gal.

\subsection{Comparison to Other Wavelength Bands}

\todo[inline]{Add comparison here to X-ray sources and to radio SNRs and to Fe II objects. ksl - It's not clear this is quite at the right place in the text. It depends a bit on whether there is anything special about the objects we have identified.}

Searches for SNRs in \gal\ have been carried out in several other wavelength ranges.  In particular,  \cite{lacey01} identified 35 radio sources as SNR candidates on the facts that these sources had non-thermal spectral indices and were positionally coincident with \ha\ emission.  As one might expect these objects are located mostly in the spiral arms of \gal\ where there is active star formation and where one might expect SNe to explode.  There are seven emission nebulae in our list of SNR candidates that lie within 2\arcsec\ of radio SNR candidates; this is five more than had been identified previously, but a small fraction of the total number of radio objects.  As an alternative to using the \sii:\ha\ ratio to identify SNRs in an external galaxy, \cite{bruursema14}  carried out ground-based interference filter imagery of \gal\ in the light of [Fe~II]1.64$\mu$.  \hii\ regions are expected to have little if any \feii, so detection of an emission nebula with \feii\ is a strong indication of shock heating.  Noting some concerns about the quality of the data,  \cite{bruursema14}  identified 48 candidate objects, only three of which, L19-076, L19-095=MF-014, and the ULX L19-098=MF-016, are in our sample of objects.

SNRs are also X-ray sources and therefore we have also looked to see how many X-ray sources in \gal\ could be found in our candidate lists.  The most detailed X-ray study of \gal\ to date was carried out using {\em Chandra} by \cite{fridriksson08}, who constructed a catalog of 90 point sources, of which 25 appeared to be time variable (and hence likely X-ray binaries).  Of the 90 point sources, there are eight which are positionally coincident with objects in our sample.

\todo[inline]{One should probably look further into the X-ray coincidences.  How many of the identified sources were time variable.  Also, MF-016 which is one of the sources is a special case. It would also be sensible to see whether the radio coincidences were particularly bright in the radio or the optical.}

%\todo[inline]{Knox: didn't you derive a more extensive X-ray source catalog yourself?  Should we be comparing to that?  (WPB)}

\subsection{Historical Supernovae in \gal}
In our 2011 WIYN emission-line images, we detected emission from only two of the nine historical SNe that had then occurred in \gal: SN\,1980K and SN\,2004et.  Both would probably have been among our \sii-selected SNR candidates, and SN\,1980K would also have attracted notice because of its relatively high \oiii:\ha\ ratio as well; however, we noted these in an explicit search of the positions of all the historical SNe in \gal.  Thus we have not included them in Tables \ref{table_candidates} or \ref{table_snr_spec}.
We  obtained GMOS spectra of both, as shown in Fig.~\ref{historical_SNe}.  Unlike the spectra from the other SNRs in our sample, the lines from both of these are highly velocity-broadened, the signature of fast-moving ejecta.

The Type IIL SN\,1980K has been frequently observed over the years since its explosion, and its transition from late-time SN to a developing SNR has been monitored both photometrically and spectroscopically \citep[e.g.,][]{uomoto86, fesen90, fesen94, fesen99b, milisavljevic12}.  The 2014 GMOS spectrum, taken 3 December 2014, about  34 yr past maximum light, is qualitatively similar to the 30 yr spectrum shown by \citet{milisavljevic12}, with broad, asymmetric lines from \ha, \oi, \oiii, and a feature near 7100 \AA\ that is probably [\ion{Fe}{2}]\,$\lambda$\,7155, possibly blended with [\ion{Ar}{3}]\,$\lambda$7136, all appearing above a faint, blue continuum \citep[see][for a discussion of the 7100 \AA\ feature]{fesen94}.  It appears that broad \SiiL\ with a similar asymmetric profile may also be present, blended with the red side of the  \ha\ line.  We estimate that the broad \ha\ flux is $8.4 \pm 1.0 \times 10^{-16} \FLUX$, slightly lower than that of $10 \pm 2 \times 10^{-16} \FLUX$ reported by \citet{milisavljevic12} for their spectrum taken in October 1010, just over 4 years earlier.  This continues the gradual decline they noted from that of $13 \pm 2 \times 10^{-16} \FLUX$ measured by \citet{fesen99b} in November 1997, which in turn represented a fading of $\sim 25\%$ from the levels observed in the early 1990s.

SN\,2004et, classified as Type IIP, was also well observed early on, and as it made the transition to its nebular stage \citep[e.g.,][]{sahu06, maguire10, jerkstrand12}; however, there appear to be no published spectra from later than 1.5 yr past its explosion.  The spectrum in Fig.~\ref{historical_SNe} shows its recovery at an age of just over 10 years.  The optical spectrum is dominated by broad, asymmetric lines of \ha, \oi, and a feature at $\sim$7150 - 7400 \AA\ that is probably a blend of (primarily)  [\ion{Fe}{2}] $\lambda 7155$ and [\ion{Ca}{2}] $\lambda\lambda 7291, 7324$, features that were prominent in its late nebular spectra \citep{sahu06, maguire10}.  (Unfortunately, this emission extends past the end of our spectrum, making it harder to identify.)  As in the case of SN\,1980K, it is possible that faint, broad \SiiL\ lines may be present, blended with the red side of \ha.  The \ha\ line flux is difficult to measure because it is so broad as to to be blended with \oi\ and, possibly, \sii; furthermore, the continuum level is also uncertain.  Our best flux estimate is $7 \pm 2 \times 10^{-16} \FLUX$.

%has had fewer late-time observations than SN\,1980K, perhaps due in part to the burgeoning number of SN discoveries over the past two decades.

% Bill added the following paragraph, but feel free to edit as necessary.
The spectra of both SN\,1980K and SN\,2004et are quite  similar to the object B12-174a identified in our similar survey of M83 \citep{blair15}. The main difference is that for B12-174a the SN was not observed, even though its inferred age is $<$100 years.  All these objects form a transitional class between ``old SNe" and mature SNRs.  SN1957D in M83 also shows broad lines, but only for oxygen, and the line intensities have dropped significantly over $\sim$40 years \citep[][and references therein]{long12}.  These differences may be due to differing progenitor types, differing local ISM conditions, or both.  \citep[The \oiii\ lines would appear relatively stronger in both  SN\,1980K and SN\,2004et if these were dereddened, with $E(B-V) = 0.41$,][but even so they would not be nearly so O-dominated as SN\,1957D.]{fesen99b, sahu06} Since there are so few objects in this transitional class, these objects bear watching for temporal changes that should happen on observable time scales.  Such observations could illuminate this poorly understood phase of SNR evolution.

The fact that none of the other seven historical SNe in \gal, ranging in age from 3 to 94 years, was detected at all is noteworthy.  All those with well-determined SN classifications are ones that result from core-collapse SNe, and thus should have produced several $M_{\sun}$ of high-velocity ejecta---the scenario responsible for ejecta-dominated SNRs like Cas A, or SN\,1957D in M83.  Furthermore, since \gal\ is such a champion producer of SNe, it is reasonable to expect the remnants from dozens of core-collapse SNe younger than 1000 yr to be located there.  The fact that so few are detected as broad-line, ejecta-dominated remnants is noteworthy.   This is similar to the case of M83, host to six historical SNe and hence also expected to have far more young SNRs than are detected.  There, \citet{winkler17} concluded that many of the SNRs are likely expanding into high-density environments, where remnants evolve rapidly to the point where they are dominated by swept-up material rather than by ejecta. At the other extreme, other SNe may have exploded in regions where earlier SNRs have evacuated the surrounding region, resulting in extremely faint SNRs.  It would seem that the situation is similar in \gal.

\section{Summary \label{sec:summary}}

We have carried out a new optical search for SNRs in \gal\ using interference filters to identify emission nebulae that have elevated \sii:\ha\ ratios compared to \hii\ regions.  Our search picked up all of the SNRs that had been identified by MF97, except one that was outside the region imaged here. Of the 147 SNR candidates we identified, we were able to obtain spectra of 102, and spectral confirmation of 89 objects.  There are 45 candidates  without spectra and 17 candidates with spectra which have spectroscopic \sii:\ha\ ratios less than the canonical ratio of 0.4 for regarding an emission nebula as a confirmed optical SNR.  Given the uncertainties, we have chosen to retain all of the 147 objects as candidates for being SNRs, though clearly only those with high ratios can be regarded as confirmed.  Very few of the candidates are detected as SNRs at other wavelengths; specifically only ?? are among the radio SNRs identified by \cite{lacey01}.  Very few (??) candidates have X-ray counterparts, which we attribute to a combination of higher absorption along the line of sight to \gal\ compared to some other galaxies at comparable distances, e.g. M83, and to the %somewhat 
lower exposure times for the {\em Chandra} study of \gal\ than for these other galaxies.

Much more work is needed to fully characterize the SNR population of \gal, some of which we are currently working on.  These include HST studies in the optical to measure diameters and identify additional small diameter objects in crowded regions, infrared \feii\ 1.64 m$\mu$ studies to identify SNRs in dusty regions or buried in complex \ha\ emission, and deeper radio studies to find and characterize the radio counterparts of the optical SNR population.

\acknowledgments

We thank NASA for allowing us time to work on this.  PFW acknowledges additional support from the NSF through grant AST-1714281.

\pagebreak

\appendix

\section{Possible additions to and notes concerning the paper
}
\todo[inline]{Frank put some questions about things one could to in the paper.  ksl has begun to delete them as they are worked off, or given as other inline comments, or if things are out of scope (e.g plots with SNR diameters for this paper.}
Here are some figures we may wish to show, and points we may want to discuss (note: KSL has done some of these figures on Evernote; those all show \oiii-selected objects, which we may choose to omit):
\begin{enumerate}
    \item Correlation between flux from spectra vs imaging (M83 paper, fig 5)  ksl - I have looked at this.  The correlation is there but quite rough.  I'm not sure it is helpful for the paper

\end{enumerate}

\pagebreak

\bibliographystyle{aasjournal}

\bibliography{snr}

\clearpage

%\todo[inline]{(Note: I moved what was Table 2 to be Table 3 so the order agrees with the text presentation order.  WPB)}

\input{table1}

\startlongtable

\clearpage

\todo[inline]{I suggest moving ``other names" to far right column, and perhaps making that more compact.  Also, check refs at bottom. -- PFW}

\input{table_candidates}

\pagebreak

\input{table3}

\clearpage

\startlongtable

\input{table_spectra}

\pagebreak
%% To help institutions obtain information on the effectiveness of their 
%% telescopes the AAS Journals has created a group of keywords for telescope 
%% facilities.
%
%% Following the acknowledgments section, use the following syntax and the
%% \facility{} or \facilities{} macros to list the keywords of facilities used 
%% in the research for the paper.  Each keyword is check against the master 
%% list during copy editing.  Individual instruments can be provided in 
%% parentheses, after the keyword, but they are not verified.

\begin{figure}
\plotone{rgb_10arcmin_subtracted_2.png}
\caption{An emission-line image of \gal, after continuum subtraction, where R = \ha, G = \sii, and B = \oiii.  The field is 10\arcmin\ square and is oriented N up, E left. \label{fig_overview} }
\end{figure}

\begin{figure}
\plotone{rgb_10arcmin_no_regions.png}
\caption{An emission-line image of \gal, where R = \ha, G = \sii, and B = \oiii.  The field is 10\arcmin\ square and is oriented N up, E left.  [Probably we will show {\bf only} one of these: before or after subtraction.] }
\end{figure}

\begin{figure}
\plottwo{fig_s2_ha_ratio_distance.pdf}{fig_spec_im_comp.pdf}
\caption{Left: The \sii:\ha\ ratio obtained spectroscopically for SNR candidates (red) and \hii\ regions (blue) as a function of galactocentric distance, R.  Objects with ratios greater than 0.4 (the dashed line) are spectroscopically confirmed SNRs.  Objects near the dividing line are less certain because observational errors in the ratio and/or \hii\ contamination can impact the derived ratio. Right: The \sii:\ha\ ratio derived from spectra compared to that derived from narrow band imaging.  The tendency for spectroscopic ratios to be somewhat higher is consistent with mild impacts from \nii\ emission getting through the \ha\ imaging filter. (This figure needs a 1:1 line to guide the eye, and remove the dashed line at 0.4 for this one--Bill) \label{fig_s2_ha}}

\end{figure}

\begin{figure}
\plotone{fig_s2_ratio.pdf}
\caption{The [S II]6717:[S II]6731 line ratio for SNR candidates as a function of galactocentric distance.  THis ratio is nominally a density diagnostic, and most candidates are near the low density limit of 1.4, indicating generally low ISM densities. Derived values above 1.4 are non-physical, and are indicative of errors in the derived ratio.  Only a handful of objects appear to show densities significantly above the low-density limit.  There is no obvious trend with galactocentric distance.  \label{fig_s2_ratio}}
\end{figure}

\begin{figure}
\plotone{fig_reddening_gc.pdf}
\caption{Observed \hb:\ha\ line ratios for SNR candidates as a function of galactocentric distance. The dashed lines indicate the expected values of the line ratio for E(B-V) of 0.0, 0.1, 0.3 and 1.0.  Objects near the galactic center tend to be more highly reddened than those far from the center. \label{fig_reddening}}
\end{figure}

\begin{figure}
\plottwo{fig_n2_ha_gc.pdf}{fig_s2_ha_gc.pdf}
\caption{Line ratios as a function of GCD.  (I assume the right panel is actually \sii:\ha?  Also, maybe show the H II regions on these plots as well?  Bill) \label{fig_metal_ratios}}

\end{figure}

\begin{figure}
\plottwo{fig_model_n2o3.pdf}{fig_model_s2n2.pdf}
\caption{Left: Model [O III] 5007:\hb\ ratio as function of the \nii\ 6583:\ha\ line ratio for SNRs and SNR candidates with spectra. As discussed in the text, the black, green and blue meshes correspond to shock models from Allen et al. (2008) with a range of shock velocities and pre-shock magnetic fields, and with metallicities corresponding to the SMC (black), LMC (green), and Milky Way (blue).  Right: Model \nii\ 6583:\ha\ line ratios as a function of the \sii:\ha\ ratios.  \gal\ objects appear consistent with solar abundances and a wide range of other physical properties.  \label{fig_model}}
\end{figure}

\begin{figure}
\plotone{spectra_historical_sne.png}
\caption{GMOS spectra of the two historical SNe we have recovered in \gal.  Both show the broad emission lines characteristic of young SNRs with rapidly expanding ejecta.  (The spectra have been displaced vertically for clarity.) \label{historical_SNe}}
\end{figure}

\vspace{5mm}
\facilities{NOAO:WIYN,
Gemini:GMOS}

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manuscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

\software{astropy \citep{astropy} 
%          Cloudy \citep{2013RMxAA..49..137F}, 
 %         SExtractor \citep{1996A&AS..117..393B}
          }

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc. The
%% Figure and Table counter will not reset.

%\appendix

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% \aastex provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.
% \nocite{}
%\begin{thebibliography}{}

%\bibliography{snr}

%\end{thebibliography}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample61.tex'.